---
title: "506_hw2"
format: html
editor: visual
---

## Problem 1

a\. Version 1

```{r}
#' @param n Number of steps.
#' @return Final position after n steps.
random_walk1 <- function(n) {
  p <- 0
  for (i in 1:n) {
    s <- sample(c(-1, 1), 1)  
    if (s == 1) {
      if (runif(1) < 0.05) {
        s <- 10}} 
    else {
      if (runif(1) < 0.20) {
        s <- -3}}
    p <- p + s}
  return(p)
}

```

a\. Version 2

```{r}
#' @param n Number of steps.
#' @return Final position after n steps.
random_walk2 <- function(n) {
  s <- sample(c(-1, 1), n, replace = TRUE)
  pm <- (s == 1) & (runif(n) < 0.05)
  s[pm] <- 10
  mm <- (s == -1) & (runif(n) < 0.20)
  s[mm] <- -3
  return(sum(s))
}

```

a\. Version 3

```{r}
#' @param n Number of steps.
#' @return Final position after n steps.
random_walk3 <- function(n) {
  st <- replicate(n, {
    s <- sample(c(-1, 1), 1)
    if (s == 1) {
      if (runif(1) < 0.05) s <- 10} 
    else {
      if (runif(1) < 0.20) s <- -3}
    s})
  return(sum(st))
}

```

Demonstration 

```{r}
set.seed(5) 
cat("After 10 steps:\n")
print(random_walk1(10))
print(random_walk2(10))
print(random_walk3(10))
cat("\nAfter 1000 steps:\n")
print(random_walk1(1000))
print(random_walk2(1000))
print(random_walk3(1000))
```

b

```{r}
stream <- function(n, seed) {
  set.seed(seed)
  d <- sample(c(-1, 1), n, replace = TRUE)   
  pr <- runif(n)                             
  list(d = d, pr = pr)
}
random_walk1 <- function(d, pr) {
  p <- 0
  for (i in seq_along(d)) {
    s <- d[i]
    if (s ==  1 && pr[i] < 0.05) s <- 10
    if (s == -1 && pr[i] < 0.20) s <- -3
    p <- p + s}
  p}
random_walk2 <- function(d, pr) {
  s <- d
  s[d ==  1 & pr < 0.05] <- 10
  s[d == -1 & pr < 0.20] <- -3
  sum(s)
}
random_walk3 <- function(d, pr) {
  s <- mapply(function(d, p) {
    if (d ==  1 && p < 0.05) return(10)
    if (d == -1 && p < 0.20) return(-3)
    d}, d, pr)
  sum(s)
}
```

Demonstration

```{r}
st <- stream(10, seed=25)
cat("For n=10:\n")
print(random_walk1(st$d, st$pr))
print(random_walk2(st$d, st$pr))
print(random_walk3(st$d, st$pr))
st <- stream(1000, seed=25)
cat("\nFor n=1000:\n")
print(random_walk1(st$d, st$pr))
print(random_walk2(st$d, st$pr))
print(random_walk3(st$d, st$pr))
```

c

```{r}
library(microbenchmark)
ben <- function(n, seed = 5, ts = 40, tl = 8) {
  st <- stream(n, seed)
  reps <- if (n < 5000) ts else tl  
  mb <- microbenchmark(
    version1 = random_walk1(st$d, st$pr),
    version2 = random_walk2(st$d, st$pr),
    version3 = random_walk3(st$d, st$pr),
    times = reps,
    unit  = "ms")
  print(mb)
  sm <- summary(mb)
  out <- sm[, c("expr", "min", "lq", "median", "uq", "max")]
  rownames(out) <- NULL
  out
}

tbs <- ben(1000, seed = 5)
tbl <- ben(100000, seed = 5)
tbs$n <- 1000
tbl$n <- 100000
co <- rbind(tbs, tbl)
names(co)[1] <- "implementation"
co <- co[, c("n", "implementation", "median", "lq", "uq", "min", "max")]
print(co)

```

Version 2 which uses the vectorized functions is the fastest in all ways. Version 3 is the slowest. Compared to the size of n = 1000, when scaling up to n = 100,000, the differences became more huge. Method of version 2 should be preferred for performance.

d

```{r}
mc <- function(n, t, batch = 1e5, seed = 5) {
  set.seed(seed)
  hits <- 0L
  dn <- 0L
  while (dn < t) {
    b <- min(batch, t - dn)
    d <- matrix(sample(c(-1L, 1L), b * n, replace = TRUE), nrow = b)
    u <- matrix(runif(b * n), nrow = b)
    s <- d
    s[d ==  1L & u < 0.05] <- 10L   
    s[d == -1L & u < 0.20] <- -3L  
    totals <- rowSums(s)
    hits <- hits + sum(totals == 0L)
    dn <- dn + b}
  p <- hits / t
  se <- sqrt(p * (1 - p) / t)
  ci <- c(p - 1.96 * se, p + 1.96 * se)
  list(p_hat = p, ci = ci, hits = hits, t = t)
}
```

```{r}
res10   <- mc(n = 10,   t = 2e6,   batch = 1e5, seed = 8)
res100  <- mc(n = 100,  t = 1e6,   batch = 1e5, seed = 8)
res1000 <- mc(n = 1000, t = 2e5,   batch = 2e4, seed = 8)
print(res10)
print(res100)
print(res1000)

```

For n = 10, the probability that the random walk ends at 0 is about 0.1321. For n = 100, the probability is about 0.0196. For n = 1000, the probability is about 0.0055.

## Problem 2

```{r}
set.seed(5)
N <- 100000  
X <- cbind(
  matrix(rpois(N * 8, 1), N, 8),                                  
  pmax(0, round(matrix(rnorm(N, 60, sqrt(12)), N, 1))),            
  matrix(rpois(N * 8, 8), N, 8),                                  
  pmax(0, round(matrix(rnorm(N, 60, sqrt(12)), N, 1))),           
  matrix(rpois(N * 6, 12), N, 6))
t <- rowSums(X)                
mc_mean <- mean(t)             
mc_se <- sd(t) / sqrt(N)     
ci <- mc_mean + c(-1, 1) * 1.96 * mc_se   
mc_mean
ci

```

So the average number of cars that pass an intersection per day under these assumptions is 264.

## Problem 3

a

```{r}
youtube <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv')
names(youtube)
```

```{r}
yc <- subset(youtube, 
        select = -c(brand, superbowl_ads_dot_com_url, youtube_url, id, kind, etag))
dim(yc)

```

b

```{r}
yc$log_view    <- log1p(yc$view_count)
yc$log_like    <- log1p(yc$like_count)
yc$log_dislike <- log1p(yc$dislike_count)
yc$log_comment <- log1p(yc$comment_count)

# Favorite count
table(yc$favorite_count)

par(mfrow=c(1,2))
# View counts
hist(yc$view_count, main="View Count", xlab="Views", col="lightblue")
hist(yc$log_view, main="View Count (log1p)", xlab="log(Views+1)", col="pink")

# Like counts
hist(yc$like_count, main="Like Count", xlab="Likes", col="lightblue")
hist(yc$log_like, main="Like Count (log1p)", xlab="log(Likes+1)", col="pink")

# Dislike counts
hist(yc$dislike_count, main="Dislike Count", xlab="Dislikes", col="lightblue")
hist(yc$log_dislike, main="Dislike Count (log1p)", xlab="log(Dislikes+1)", col="pink")

# Comment counts
hist(yc$comment_count, main="Comment Count", xlab="Comments", col="lightblue")
hist(yc$log_comment, main="Comment Count (log1p)", xlab="log(Comments+1)", col="pink")

```

-   View counts. ii) The variable can use a transformation prior to being used as the outcome in a linear regression model.

-   Like counts.  ii) The variable can use a transformation prior to being used as the outcome in a linear regression model.

-   Dislike counts. ii) The variable can use a transformation prior to being used as the outcome in a linear regression model.

-   Favorite counts. iii) The variable would not be appropriate to use as the outcome in a linear regression model.

-   Comment counts. ii) The variable can use a transformation prior to being used as the outcome in a linear regression model.

c

```{r}
library(broom)    
library(dplyr)    
library(knitr)      
rview    <- lm(log_view    ~ funny + show_product_quickly + patriotic +
                  celebrity + danger + animals + use_sex + year, data = yc)
rlike    <- lm(log_like    ~ funny + show_product_quickly + patriotic +
                  celebrity + danger + animals + use_sex + year, data = yc)
rdislike <- lm(log_dislike ~ funny + show_product_quickly + patriotic +
                  celebrity + danger + animals + use_sex + year, data = yc)
rcomment <- lm(log_comment ~ funny + show_product_quickly + patriotic +
                  celebrity + danger + animals + use_sex + year, data = yc)
results <- list(
  View    = rview,
  Like    = rlike,
  Dislike = rdislike,
  Comment = rcomment) %>%
  lapply(tidy) %>%
  bind_rows(.id = "Outcome") %>%
  filter(term != "(Intercept)") %>%
  mutate(significant = ifelse(p.value < 0.05, "Yes", ""))

kable(results[, c("Outcome", "term", "estimate", "std.error", "p.value", "significant")],
      digits = 3, caption = "Linear regression")


```

It shows that year is a statistically significant positive predictor for likes and dislikes. And none of the characteristics such as funny, celebrity and animals showed robust significant effects.

d

```{r}
form <- log_view ~ funny + show_product_quickly + patriotic + celebrity + danger + animals + use_sex + year
mf <- model.frame(form, data = yc, na.action = na.omit)
X <- model.matrix(form, data = mf)    
y <- model.response(mf)
beta_hat_m <- qr.solve(X, y)

fit_lm <- lm(form, data = yc)       
beta_lm <- coef(fit_lm)

cat("Manual beta hat:\n"); print(beta_hat_m)
cat("\nlm coefficients:\n"); print(beta_lm)
cat("\nEqual? ", all.equal(as.numeric(beta_hat_m), as.numeric(beta_lm), tolerance = 1e-10), "\n", sep = "")


```
